Titeln på avhandlingens språk

### 2023-04-24 body/part1/chapter1-introduction/introduction.tex
introduction

\todo[inline]{
    Needs to be reworked based on scope, new formulation of two contributions.
    Enumerate contributions at the end.
}

@@

These limitations of the Cloud make it unsuitable for a number of novel applications that require quick response times, such as real-time video processing, \gls{MAR}, or smart cities.
Additionally, the centralized nature of Cloud datacenters means that it can be vulnerable to outages or other disruptions, which can impact large numbers of users simultaneously.

@@

By placing computing resources in proximity to where data is generated and/or consumed, and thus reducing the topological and physical distances that data must travel, Edge Computing can greatly decrease latency, improve the responsiveness of applications, and allow for greatly improved efficiency in the use of network resources.
This enables powerful and sophisticated applications to be run on the Edge, such as \glspl{CPS} as well as \gls{XR} applications.
These applications have the potential to have a profound impact in our day-to-day lives, but have stringent latency requirements which make their deployment on traditional paradigms such as Cloud Computing inviable.

@@

Overall, the combination of Edge Computing and 5G holds great promise for the massive deployment of latency-sensitive applications such as the ones discussed above.
Nonetheless, significant challenges remain in the understanding and scaling of these systems, particularly in regard to their complexity and requirements in the context of low-latency, high-reliability computing.
Before widespread deployment of these systems, it is crucial to thoroughly understand their real-world performance.

\glspl{CPS} often correspond to safety-critical components of industrial systems, and subpar \gls{QoS} can therefore lead to significant danger to equipment and operators.
On the other hand, the immersive nature of \gls{XR} applications implies that poor \gls{QoS} can easily translate into significant user discomfort.
To ensure reliability, safety, and user satisfaction in these systems, careful and accurate characterization and optimization of these systems must be realized.

@@

The evaluation of performance in Edge computing systems has been addressed in existing literature through various approaches such as analytical, simulation, and experimental methods.
However, the unique nature of Edge computing environments demands a wide range of interdisciplinary expertise spanning (but not limited to) computer science, communication networks, information theory, queuing analysis, etc.
Such expertise may not always be feasible to obtain within a single research group.
This has led to prior research to often be constrained in its scope, focusing on comprehensive characterization of either compute or network aspects, resulting in potential trade-offs that may neglect one half of the larger picture.
We argue that this limitation has led to an incomplete understanding of the overall performance of Edge systems and applications in the existing literature.
There is a need for holistic approaches that consider both network and compute characteristics of Edge workloads for accurate performance evaluation and optimization.

At the time of the writing of this dissertation, no widely adopted methodology exists for the study of these trade-offs.
We aim to contribute to the field by introducing and subsequently investigating the applications of such a methodology for the study of latency-sensitive applications deployed on Edge infrastructure and network technologies such as 5G.
The proposed methodology aims to enhance the accuracy and realism of results related to Edge infrastructure and 5G networks, particularly in regard to network performance.
Our results will contribute to the development of new techniques and approaches for improving the performance and reliability of \gls{MEC}.

One of the main challenges to the study of latency-sensitive applications such as \gls{CPS} and \gls{MAR} on Edge Computing is that these applications interact with the real world, and thus must be studied in real-time.
This implies that any study of these applications must be done in an environment that accurately reflects the real-world conditions in which they are deployed.
Unfortunately, replicating these conditions in a controlled experimental environment is difficult and often leads to inconsistencies in results, and traditional methods that rely on simulations may not be sufficient to accurately capture and evaluate the performance of these applications in real-world scenarios.

The scalability of studying these systems is also a challenge.
It may not be feasible for researchers to test the applications on a large scale due to limitations in access to specialized equipment, personnel and test subjects, as well as facilities.
\glspl{CPS} often rely on physical sensors and actuators, making it difficult to replicate real-world conditions in a controlled environment at scale, and conducting experiments involving human subjects for \gls{MAR} is time-consuming and expensive.

The highly complex effects on the wireless network is another major challenge to studying latency-sensitive applications on Edge Computing.
The presence of multiple devices accessing the network simultaneously can result in interference, network congestion, and other factors that can affect the performance of the system.
These effects are difficult to simulate and/or emulate, particularly in multi-tenant deployments, where each tenant may have different requirements and resource usage patterns.

Finally, the combination of the characteristics enumerated above makes the study of these applications hard to automate, making it challenging to conduct large-scale experiments or to collect data over an extended period of time.
\gls{CPS} and \gls{MAR} applications are highly dynamic and adaptive, and their behavior cannot be easily predicted.
Moreover, there is often a significant amount of variability in the behavior of different components of the application, such as sensors, actuators, and network connections, which makes it difficult to automate the testing process.

Our methodology is designed to tackle these challenges by employing an emulation approach to the benchmarking of latency-sensitive \gls{CPS} and \gls{MAR} applications. 
We emulate target workloads on actual Edge infrastructure by replacing the client side of the system with \gls{COTS} general-purpose computing devices\footnote{%
    In our initial implementation, these correspond to low-cost and easily replaceable, scalable Raspberry Pi 4 Model B \acsp{SBC}.
} running a realistic software imitation of the desired behaviors.
At the same time, our methodology retains the real network as well as the real compute hardware and software at the backend.
This design results in real-time benchmarking of applications, while still providing realism, scalability, and reduced complexity, as well as allowing for easier automation of these studies.

Emulating the workload component reduces complexity by moving it into the software domain, allowing for easier horizontal scaling through the use of \gls{COTS} general-purpose hardware such as \glspl{SBC}.
This also reduces the barrier of entry to this research, as \glspl{SBC} are for the most part cheap and easily accessible.
Additionally, it preserves the realism of effects stemming from the hardware and network.
In particular, the methodology allows us to capture effects due to network factors such as contention, congestion control, and medium access, which are often of stochastic or chaotic natures and complex to capture in simulations.

The methodology also provides automation as well as improved repeatability and replicability.
The softwarized nature of the client-side emulation allows for straightforward automation through general-purpose programming and scripting languages.
Repeating a study becomes a matter of re-running the workload on the same testbed, and studies can be replicated simply by obtaining the same or equivalent software workload and deploying it on a comparable testbed.
These are complex tasks to accomplish in real-world approaches, particularly when dealing with humans.

We validate our proposed methodology by presenting case studies of its application in two different contexts: \gls{WCA} and \glspl{NCS}.
The case studies demonstrate the efficacy of the methodology in evaluating system performance and overall \gls{QoS} and \gls{QoE}.
Our methodology provides a comprehensive and realistic assessment of the performance of latency-sensitive applications deployed on Edge infrastructure.
The approach provides valuable insights for researchers, system designers, and application developers, and will contribute to the development of new techniques and approaches for improving the performance and reliability of Edge Computing and 5G systems.

In addition to the methodology, this dissertation introduces a number of further contributions to the field. 
We introduce a software framework that facilitates the orchestration of Edge Computing testbeds needed for the implementation of our methodology.
This framework makes it easier for researchers and practitioners to deploy and test Edge-bound low-latency applications.
We also present a deep exploration of this methodology for \gls{WCA} and introduce the first ever model for the end-to-end emulation of human timing behavior in \gls{WCA}.
This model is built from a thorough characterization of these behaviors, the data for which we obtain from a comprehensive human-subject study.

Finally, we explore the implications of our methodology and human user model on the optimization potential of \gls{WCA} deployments.
In particular, we study the sampling and energy consumption behaviors of these systems with and without considering realistic human behavior.
We conclude that significant improvements can be achieved through the use of our human user model.
Overall, our results showcase the utility of our methodology, and the importance of considering human behavior in the design and optimization of \gls{WCA} systems to achieve maximum benefits.

>>>>

@@

These limitations of the Cloud make it unsuitable for a number of novel applications that require quick response times, such as real-time video processing, \gls{MAR}, or smart city systems.
Additionally, the centralized nature of Cloud datacenters means that it can be vulnerable to outages or other disruptions, which can impact large numbers of users simultaneously.

@@

By placing computing resources in proximity to where data is generated and/or consumed, and thus reducing the topological and physical distances that data must travel, Edge Computing can greatly decrease latency, improve the responsiveness of applications, and allow for greatly improved efficiency in the use of network resources.
This enables powerful and sophisticated applications to be run on the Edge, such as \glspl{CPS} as well as \gls{XR} applications.
These applications have the potential to cause a profound impact in our day-to-day lives, but have stringent latency requirements which make their deployment on traditional paradigms such as Cloud Computing inviable.

@@

Overall, the combination of Edge Computing and 5G holds great promise for the massive deployment of latency-sensitive applications such as the ones discussed above.
Nonetheless, significant challenges remain in the understanding and scaling of these systems, particularly in regard to their complexity and requirements in the context of low-latency, high-reliability computing.

Before widespread deployment of these systems, it is crucial to thoroughly understand their real-world performance.
\glspl{CPS} often correspond to safety-critical components of industrial systems, and subpar \gls{QoS} can therefore lead to significant danger to equipment and operators.
On the other hand, the immersive nature of \gls{XR} applications implies that poor \gls{QoS} can easily translate into significant user discomfort.
To ensure reliability, safety, and user satisfaction in these systems, careful and accurate characterization and optimization of these systems must be realized.

@@

The evaluation of performance in Edge computing systems has been addressed in existing literature through analytical, simulation, and experimental approaches.
However, these systems exhibit several characteristic that make their study complex.
One of these is unique nature of Edge computing environments, which tightly integrates compute and network.
This demands a wide range of interdisciplinary expertise to study, spanning (but not limited to) computer science, communication networks, information theory, queuing analysis, etc.
Such expertise may not always be feasible to obtain within a single research group.

The highly complex effects emerging from the interaction between network and compute is another major hurdle to studying latency-sensitive applications on Edge Computing.
These effects are specially difficult to address through analytical models and simulations, particularly in multi-tenant deployments, where each tenant may have different requirements and resource usage patterns.
Although powerful approaches to the modeling of network effects and compute process scheduling exist in literature, existing analytical and simulation approaches have so far not managed to combine the two.

Nevertheless, experimental approaches also face important challenges.
One of these stems from the fact that latency-sensitive applications such as \gls{CPS} and \gls{XR} on Edge Computing interact with the real world.
This implies that these applications thus must be studied in real-time to achieve comprehensive characterization.
Thus, any study of these applications must be done in an environment that accurately reflects the real-world conditions in which they are deployed.
Unfortunately, replicating these conditions in a controlled experimental environment is difficult and often leads to inconsistencies in results, and traditional methods that rely on simulations may not be sufficient to accurately capture and evaluate the performance of these applications in real-world scenarios.

The scalability of these systems is also a challenge.
It may not be feasible for researchers to test the applications on a large scale due to limitations in access to specialized equipment, personnel and test subjects, as well as facilities.
\glspl{CPS} often rely on physical sensors and actuators, making it difficult to replicate real-world conditions in a controlled environment at scale, and conducting experiments involving human subjects for \gls{XR} is time-consuming and expensive.
This also makes the study of these applications hard to automate, making it challenging to conduct large-scale experiments or to collect data over an extended period of time.

This has led to existing approaches to performance evaluation of Edge Computing environments to often be constrained in their scope, focusing on comprehensive characterization of either compute or network aspects, resulting in potential trade-offs that may neglect one half of the larger picture.
We argue that this limitation has led to an incomplete understanding of the overall performance of Edge systems and applications in the existing literature.
There is a need for holistic approaches that consider both network and compute characteristics of Edge workloads for accurate performance evaluation and optimization.

This dissertation aims to contribute to the field by presenting two key contributions.
First, we introduce and subsequently investigate the applications of a methodology for the study of latency-sensitive applications deployed on Edge environments.
The proposed methodology is based on the emulation of workloads and aims to enhance the accuracy and realism of results related to Edge infrastructure and 5G networks, particularly in regard to network performance.
Next, we present a deeper exploration of the implications of improved realism in the emulated components of our methodology by studying potential avenues for optimization of a particular sub-class of \gls{MAR}.

Our methodology is designed to tackle these challenges by employing an emulation approach to the benchmarking of latency-sensitive \gls{CPS} and \gls{XR} applications. 
We emulate target workloads on actual Edge infrastructure by replacing the client side of the system with \gls{COTS} general-purpose computing devices\footnote{%
    In our initial implementation, these correspond to low-cost and easily replaceable, scalable Raspberry Pi 4 Model B \acsp{SBC}.
} running a realistic software imitation of the desired behaviors.
At the same time, our methodology retains the real network as well as the real compute hardware and software at the backend.
This design results in real-time benchmarking and performance evaluation of applications and Edge infrastructure, while still providing realism, scalability, and reduced complexity, as well as allowing for easier automation of these studies.

Emulating the workload component reduces complexity by moving it into the software domain, allowing for improved repeatability and replicability, as well as easier horizontal scaling through the use of \gls{COTS} general-purpose hardware such as \glspl{SBC}.
This also reduces the barrier of entry to this research, as \glspl{SBC} are for the most part cheap and easily accessible.
Additionally, it preserves the realism of effects stemming from the hardware and network.
In particular, the methodology allows us to capture effects due to the interaction of network factors such as contention, congestion control, and medium access, with effects arising due to process scheduling and/or resource contention at the compute level.
These effects are often of stochastic or chaotic natures and complex to capture in models or simulations.

Next, we study how improved realism in the emulations employed by this methodology can open new avenues for the optimization of applications on Edge Computing.
To achieve this, we develop a novel model for human behavior in a subclass of \gls{MAR} which is able to dynamically adapt to changes in system responsiveness.
We then combine it with a mathematical framework to study the potential for optimization of resource consumption in these applications.
We find that improved realism in the emulation enables otherwise inviable optimization approaches.

We believe our contributions represent valuable tools and insights for researchers, system designers, and application developers.
The methodology represents a novel, holistic approach to the performance evaluation of Edge environments.
Furthermore, our findings have significant implications for the design and optimization of applications enabled by Edge Computing, particularly in the areas of energy consumption, sampling strategies, and application lifetime.
We hope and believe this dissertation will contribute to the development of new techniques and approaches for improving the performance and reliability of Edge Computing and 5G systems.
